The Form of Uncertainty Affects Selection for Social Learning

### 2023-02-21 ms.tex
reversed order in Figure 5
revisions nearly done

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several forms of uncertainty and endows agents
with a relatively powerful mechanism for individual learning. We model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection-set size (the number of possible behaviors); (3) payoff ambiguity
(the difference in the expected rewards for different behaviors); and (4) the effective lifespan of agents (the number of
opportunities for individual learning), see next section for more details. In our model, social learners learn from the previous
generation and asocial learners do not. All agents engage in individual learning
within their lifespans. The model also assumes \edit{an adaptive},
empirically-motivated learning mechanism, the softmax decision rule, which allows learners to make full use of the available data acquired by both individual and social learning \citep{SuttonBartoBook}. Our model design thereby affords us a richer and possibly more accurate assessment of trade-offs between social and asocial learning. We use this integrated model to ask which kinds of uncertainty are likely to favor social learning and to clarify the logic underlying the results.
In the remainder of this introduction we briefly review previous research on both the forms of uncertainty and the learning mechanism used in our model.

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values (Table~\ref{tab:modelParameters}) and observed how frequently social learning evolved across 1000 trial simulations for each combination of parameters tested. %We varied $u \in \{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  
We observed three outcome measures, with $\meansl$ as the primary measure: the average value of $s_i$ over all agents and trials. To support our conclusions about $\meansl$, we also measured $\meanG$, the mean number of generations to fixation, and $\meanpi / L$, the average net  payoffs at the end of the simulation across agents and trials, normalized by lifespan (see Table~\ref{tab:outcomeVariables}). We analyze
outcomes by plotting $\meansl$ on the y-axis and environmental variability on
x-axis, since we theoretically expect that $\meansl$ will always decrease monotonically from 1 to 0 as $u$ increases. %Note that the hypothesis-testing concept of significance is meaningless here because we could make any small outcome difference ``significant'' by running more simulation trials.  We instead study the patterns of outcome variables over systematically varied uncertainty factor values.

\begin{table}[h]
    \caption{Outcome Variables. All averages are computed across trials at the end of the last generation.}
    \label{tab:outcomeVariables}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.25in}p{0.85in}} \toprule

        Symbol & Description & Values \\

        $\meansl$ & Mean social-learning prevalence over agents
                  & $ [0.0, 1.0]$ \\

        $\meanpi / L$ & Mean payoffs accumulated across agents, normalized by
        lifespan & $ [0.0, 1.0]$ \\

        $\meanG$ & Mean number of generations to convergence & \{1,2,\ldots\} \\
        \bottomrule
    \end{tabular}
\end{table}

@@

Our main results are illustrated by Figure~\ref{fig:mainResults}, which shows the
proportion of simulation runs that fixated at 100\% social learners as a function of
each of our four uncertainty measures. First, we found that reliance on social
learning monotonically decreases as temporal environmental variability, $u$,
increases. When the environment is more likely to change between generations,
information learned from the previous generation is less likely to be of value,
decreasing selection for social learning. Past some threshold, social learning is
not favored at all. However, the exact nature of the relationship between social
learning and temporal environmental variability was moderated by the other three uncertainty parameters.

The number of behavioral options, i.e., larger $B$, favors the evolution of social learning. Increasing the selection-set size
expands the range of $u$ over which social learning is favored for two reasons. First, when the environment remains constant between
generations, the value of social learning increases with more behavioral options, because it effectively increases the number of
observations that one can learn from by comparing the payoffs of several models. In other words, relying \emph{only} on one's
individual trial-and-error learning is less likely to yield useful information when the search space is large. Second, when the environment changes, bias against the new optimal is weaker with more behavioral options. For example, with only two behavioral choices ($B = 2$), a common assumption in many models, social learning is particularly detrimental when the environment changes since more of the population will have arrived at the optimal behavior in the previous generation and socially learning from them will be biased against the correct choice. These two mechanisms make it particularly profitable to engage in social learning when the selection-set size is larger---for large values of $B$ social learning can even be favored when the environment is more likely to change than not (i.e., $u>0.5$).

A shorter effective lifespan, $L$, also favors the evolution of social learning in
most cases. When effective lifespans are long and there are many opportunities to
gather information by individual learning, the value of social transmission is
diminished. Given enough opportunities, all agents will learn the optimal behavior
eventually. Thus, under most conditions, longer effective lifespans lead to a
narrower range of the parameters $u$ and $B$ under which social learning was
favored. It is noteworthy that in many cases, a large increase in the selection
regime for social learning is observed when $L=1$, and by definition no individual
learning can occur. The exceptions to this general pattern occur for high levels of
payoff ambiguity, $\pilow$ combined with low rates of environmental change, $u$
(left parts of plots on bottom row of Figure~\ref{fig:mainResults}). This reversal
is partially explained by \edit{weaker selection} when there is little difference in the payoffs between behaviors, and by the fact that the quality of social information is poorer when payoffs are ambiguous and lifespans are short (more on these effects of payoff ambiguity below).  
% \cm{is this right? I thought we should say something about the exceptions in this paragraph though we describe in more detail below, but not sure I have the right explanation}

@@

\begin{figure}
  \caption{Social learning prevalence ($y$-axes) monotonically decreases as 
  environmental variability, $u$, increases ($x$-axes) in most uncertainty contexts. 
  Other uncertainty values payoff ambiguity, $\pilow$, (rows), number of behavioral options, $B$, (columns), and effective lifespan, $L$, (keys) shift and flatten the slope from all-social-learner populations to all-asocial-learner
  populations.}
  \label{fig:mainResults}
  \centering
    \includegraphics[width=\textwidth]{Figures/mainResultsPlots.pdf}
\end{figure}

@@

Figure~\ref{fig:payoffs} shows the average net payoffs normalized by lifespan
for our main simulation compared with the two reference homogeneous populations
initialized either as all-social or all-asocial learners. Often the simulated payoffs follow the payoffs of the better-performing homogeneous group. However, there are several important deviations worth explaining. First, agents from our main simulations who evolved to be social learners sometimes outperformed agents in the homogeneous social learner populations. This is most pronounced when $L=1$, and even more
so when $\pilow=0.45$ and $B=10$ (Figure~\ref{fig:payoffs}, middle right).
In fact, social learning evolved even in parts of the parameter space where the expected homogeneous social-learning payoff was lower than the expected homogeneous
asocial-learning payoff (Figure~\ref{fig:payoffs}, e.g., middle center $L=1$).
Second, we also observed the opposite dynamic: populations sometimes evolved to be asocial learners, even though all-social learner populations outperformed all-individual-learner populations when agents were long-lived ($L=4$ and $L=8$) and in simple environments ($B=2$; Figure~\ref{fig:payoffs}, e.g., upper left).

@@

When lifespan was minimal ($L=1$) and environmental variability was moderate ($u\approx0.5$), the payoff time series showed a period
of environmental stability that
could lead the population to fixate to social learning, despite individual learning being the superior long-term strategy (Figure~\ref{fig:payoffTimeseries}a).
On the other hand, when payoff ambiguity and selection-set size was small ($\pilow=0.1$ and $B=2$), and lifespan was long ($L \geq 4$), asocial learning fixated as a hedge against periods of protracted environmental instability, even though expected
homogeneous social payoffs were greater than homogeneous asocial payoffs (Figure~\ref{fig:payoffTimeseries}b). Runs of environmental change strongly affected the geometric mean fitness of social learners who were weighing
outdated information too heavily. %\cm{is there something worth pointing out as different between b and c? If not maybe just choose one? I'd choose b}

\begin{figure}
  \caption{Average number of generations ($\meanG$, y-axes) to fixation. 
    Note the y-axis scale varies between plots, indicating different overall 
    time to fixation for different $(\pilow, B)$ combinations.} 
  \label{fig:steps}
\centering
    \includegraphics[width=\textwidth]{Figures/stepResultsPlots.pdf}
\end{figure}

@@

\begin{figure}
    \caption{Example time series of geometric moving average (GMA) with window of 3
      of payoffs from two select uncertainty settings (see main text), broken out by
    social learners, asocial learners, and whole population, compared with the
    expected homogeneous social and asocial population payoffs. Social-learner prevalence is also plotted. Vertical lines indicate environmental change.}
    \label{fig:payoffTimeseries}
  \centering

\vspace{1em}
\hspace{-2em}
    \begin{subfigure}[]{0.75\textwidth}
      \centering
    \caption{$u=0.5$, $\pilow=0.1$, $B=4$, $L=1$}
      \includegraphics[width=0.95\textwidth]{Figures/geopayoff_series/geopayseries_u=0.5-lowpayoff=0.1-nbehaviors=4-L=1.pdf}
    \end{subfigure} \\

\hspace{-2em}
  \begin{subfigure}[]{0.75\textwidth}
      \centering
      \caption{$u=0.2$, $\pilow=0.1$, $B=2$, $L=4$}
      \includegraphics[width=0.95\textwidth]{Figures/geopayoff_series/geopayseries_u=0.2-lowpayoff=0.1-nbehaviors=2-L=4.pdf}
    \end{subfigure} %\\

@@

 Despite the common claim that copying others makes sense when one is uncertain, we find that different forms of uncertainty can
 either promote or impede the evolution of social learning. By disambiguating various formalizations of uncertainty in a single
 computational model, we have developed a more nuanced theoretical understanding of how uncertainty affects the evolution of social learning. %(but see \citet{healy2017}).
%We reviewed and distilled several empirical and theoretical models in the literature along four uncertainty dimensions: temporal environmental uncertainty, selection-set size, ambiguous payoffs, and effective lifespan.  
We reproduced the well-known results that the uncertainty derived from a temporally-variable environment can limit the evolution of social learning if environmental change is simply too frequent for intergenerational transmission to be of value. Importantly, we also showed that this relationship is affected by other sources of uncertainty. 
%Our model reproduced classic predictions that social learning can yield higher payoffs than individual learning when uncertainty is sufficiently limited. 
We found that larger sets of behavioral options, shorter effective lifespans,
and increased payoff ambiguity could all favor the evolution of social learning.
However, these effects were not always straightforward. Increased payoff
ambiguity and decreased effective lifespan weakened selection strength
(i.e., evolutionary uncertainty in a finite-size population). Not surprisingly,
when the consequence of one's choice does not affect payoffs much, selection
either for or against social learning is substantially weaker. More
interestingly, selection is also weaker when the ability to learn individually
is fraught with uncertainty. Our model demonstrated that social learning can act
as a scaffold for \edit{adaptive} individual learning. We endowed agents with a
biologically-plausible individual learning mechanism, softmax search. This
enabled agents to recover from the potential cost of receiving outdated
information. This, in turn, enabled the evolution of social learning even when
social information was likely to be outdated.

@@

Our model necessarily made simplifying assumptions. These were chosen judiciously to
address the main question of how uncertainty affects the evolution of social
learning. Two critical assumptions we made have important consequences for comparing
our results with previous findings in the literature. First, we assumed all agents
engage in individual learning. This means that individual learning is effectively
free, though potentially inaccurate. Although producing new knowledge may be costly,
many other adaptive problems take the form of essentially cost-free individual
learning. For example, it is hard not to encode association between clouds and rain,
between thirst and drinking water, or between friendship and warm laughter.
\edit{Our model operationalizes environmental variability as fully stochastic,
which forces evolution towards either a population of social or asocial learners
since environmental change is unpredictable, unlike precursor
models that allowed for persistent frequency
dependence~\citep{BoydRicherson1985,Rogers1988,Feldman1996}.} Second, we only allow for
intergenerational social learning and environmental change. Within generations,
agents can only learn individually. This assumption allows us to replicate the
classic finding that if the environment changes too quickly, socially-learned
information may become \edit{outdated. Adding horizontal social learning between
vertical/oblique learning in this model would likely expand the range of
environmental variability over which social learning would
evolve~\citep{Turner2022}, i.e., horizontal learning would increase what we have
called the social learning extinction variability. However, depending on the
setting, there may be little difference between horizontal and oblique learning, for
example if the generations are purely cultural ‘generations’ where parents, and
others of the parent's generation, may also be peers.}

@@

%We found that the 
The evolution of social learning is a complex phenomenon, dependent on many interrelated factors, yet it is widespread across a
range of taxa. Although social learning is broadly studied, few studies have taken a systematic approach to understanding the role
of uncertainty in the evolution \edit{of} social learning. Understanding the role of uncertainty in social evolution is particularly important
for understanding human evolution as evidence increasingly suggests that the genus \emph{Homo} emerged in the context of substantial
(nonstationary) environmental variability \citep{anton_etal2014, levin2015}, and our global dispersal suggests that human groups regularly face novel environments. By carefully identifying and operationalizing different forms of uncertainty common in social-learning models, we developed a more thorough understanding of how uncertainty can favor social learning and how different forms of uncertainty interact to favor its evolution. By explicitly modeling a key individual-level learning mechanism shared across taxa, namely the ability to adjust behavior to uncertainty and new information, we found that social learning could evolve even if it provides outdated information. %Social learning is a critical component of problem solving among social creatures. Our work identified and systematized some important forms of uncertainty and catalogued their effects on the evolution of social learning.
Our work, then, both advances the theory of social learning and could eventually be of practical use in identifying specific mechanisms underlying adaptations to uncertain environments.

@@

\paragraph{Author contributions.} JHJ, CM, and PES conceived of the project. MAT led model development and analysis, with contributions from CM, PES, and JHJ. MAT wrote and edited the initial manuscript draft, which was extensively expanded and edited by CM, PES, and JHJ.

>>>>

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several forms of uncertainty and endows agents
with a relatively powerful mechanism for individual learning. We model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection-set size (the number of possible behaviors); (3) payoff ambiguity
(the difference in the expected rewards for different behaviors); and (4) the effective lifespan of agents (the number of
opportunities for individual learning), see next section for more details. In our model, social learners learn from the previous
generation and asocial learners do not. All agents engage in individual learning
within their lifespans. The model also assumes \edit{an adaptive},
empirically-motivated learning mechanism, the softmax \edit{decision rule}, which allows learners to make full use of the available data acquired by both individual and social learning \citep{SuttonBartoBook}. Our model design thereby affords us a richer and possibly more accurate assessment of trade-offs between social and asocial learning. We use this integrated model to ask which kinds of uncertainty are likely to favor social learning and to clarify the logic underlying the results.
In the remainder of this introduction we briefly review previous research on both the forms of uncertainty and the learning mechanism used in our model.

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values %(Table~\ref{tab:modelParameters}) 
and observed how frequently social learning evolved across 1000 trial simulations for each combination of parameters tested. %We varied $u \in \{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  
\edit{We observed three outcome measures: first, the social learning (SL) fixation ferquency, denoted
$\meansl$, is how frequently
social learning evolved to fixation across all trials for a given parameter setting. To support our
analysis about why we observe the patterns in social learning fixation frequency that we did, we also
measured the mean number of generations to fixation, which measures the strength of selection.} Finally we
measured the average net payoffs at the end of the simulation across agents and trials, normalized by lifespan. We analyze
outcomes by plotting $\meansl$ on the y-axis and environmental variability on
x-axis, since we theoretically expect that $\meansl$ will always decrease monotonically from 1 to 0 as $u$ increases. %Note that the hypothesis-testing concept of significance is meaningless here because we could make any small outcome difference ``significant'' by running more simulation trials.  We instead study the patterns of outcome variables over systematically varied uncertainty factor values.

% \begin{table}[h]
%     \caption{Outcome Variables. All averages are computed across trials at the end of the last generation.}
%     \label{tab:outcomeVariables}
%     \centering %\hspace{-3em}
%     \begin{tabular}{cp{4.25in}p{0.85in}} \toprule

%         Symbol & Description & Values \\

%         \midrule

%         $\meansl$ & Mean social-learning prevalence over agents
%                   & $ [0.0, 1.0]$ \\

%         $\meanpi / L$ & Mean payoffs accumulated across agents, normalized by
%         lifespan & $ [0.0, 1.0]$ \\

%         $\meanG$ & Mean number of generations to convergence & \{1,2,\ldots\} \\
%         \bottomrule
%     \end{tabular}
% \end{table}

@@

Our main results are illustrated by Figure~\ref{fig:mainResults}, which shows the
proportion of simulation runs that fixated at 100\% social learners as a function of
each of our four uncertainty measures. First, we found that reliance on social
learning monotonically decreases as temporal environmental variability, $u$,
increases. When the environment is more likely to change between generations,
information learned from the previous generation is less likely to be of value,
decreasing selection for social learning. \edit{Past some threshold, social learning never evolves.  We define this threshold environmental variability as the \emph{social learning extinction variability} (Figure~\ref{fig:extinctionVariability}).} 
However, the exact nature of the relationship between social
learning and temporal environmental variability was moderated by the other three uncertainty parameters.

The number of behavioral options, i.e., larger $B$, favors the evolution of social learning. Increasing the selection-set size
expands the range of $u$ over which social learning is favored for two reasons. First, when the environment remains constant between
generations, the value of social learning increases with more behavioral options, because it effectively increases the number of
observations that one can learn from by comparing the payoffs of several models. In other words, relying \emph{only} on one's
individual trial-and-error learning is less likely to yield useful information when the search space is
large. Second, when the environment changes, bias against the new optimal is weaker with more behavioral
options. For example, with only two behavioral choices ($B = 2$), a common assumption in many models, social
learning is particularly detrimental when the environment changes since more of the population will have
arrived at the optimal behavior in the previous generation and socially learning from them will be biased
against the correct choice. These two mechanisms make it particularly profitable to engage in social
learning when the selection-set size is larger---for large values of $B$ social learning can even be favored
when the environment is more likely to change than not (i.e.,
$u>0.5$\edit{---see Figure~\ref{fig:extinctionVariability}}).

A shorter effective lifespan, $L$, also favors the evolution of social learning in
most cases. When effective lifespans are long and there are many opportunities to
gather information by individual learning, the value of social transmission is
diminished. Given enough opportunities, all agents will learn the optimal behavior
eventually. Thus, under most conditions, longer effective lifespans lead to a
narrower range of the parameters $u$ and $B$ under which social learning was
favored. It is noteworthy that in many cases, a large increase in the selection
regime for social learning is observed when $L=1$, and by definition no individual
learning can occur. The exceptions to this general pattern occur for high levels of
payoff ambiguity, $\pilow$ combined with low rates of environmental change, $u$
(left parts of plots on bottom row of Figure~\ref{fig:mainResults}). This reversal
is partially explained by \edit{weaker selection} when there is little difference in the payoffs between
behaviors, and by the fact that the quality of social information is poorer when payoffs are ambiguous and
lifespans are short (more on these effects of payoff ambiguity below).  
\edit{When $L$ becomes roughly equal to or greater than $B$ there is little marginal effect of increasing $L$, indicating that more individual learning experiences are unlikely to significantly affect an agent's ability to identify the optimal behavior in their lifetime.}
% \cm{is this right? I thought we should say something about the exceptions in this paragraph though we describe in more detail below, but not sure I have the right explanation}

@@

\begin{figure}
  \caption{Social learning fixation frequency across trials (y-axes) monotonically decreases as 
  environmental variability, $u$, increases (x-axes) in most uncertainty contexts. 
  Other uncertainty values payoff ambiguity, $\pilow$, (rows), number of behavioral options, $B$, (columns), and effective lifespan, $L$, (keys) shift and flatten the slope from all-social-learner populations to all-asocial-learner
  populations.}
  \label{fig:mainResults}
  \centering
    \includegraphics[width=\textwidth]{Figures/mainResultsPlots.pdf}
\end{figure}

\begin{figure}
  \caption{Social learning goes extinct for larger values of environmental variability (SL extinction
    varibility, y-axis) as the number of
  behaviors afforded by the environment increases (x-axis).}
  \label{fig:extinctionVariability}
  \centering
    \includegraphics[width=0.8\textwidth]{Figures/SL_Extinction_Variability.pdf}
\end{figure}

@@

Figure~\ref{fig:payoffs} shows the average net payoffs normalized by lifespan
for our main simulation compared with the two reference homogeneous populations
initialized either as all-social or all-asocial learners. Often the simulated payoffs follow the payoffs of the better-performing homogeneous group. However, there are several important deviations worth explaining. First, populations sometimes evolved to be asocial learners, even
though all-social learner populations outperformed all-individual-learner populations when agents were
long-lived ($L=4$ and $L=8$) and in simple environments ($B=2$; e.g., Figure~\ref{fig:payoffs}A).
Second, we also observed the opposite dynamic: agents from our main simulations who evolved to be social learners sometimes outperformed agents in the homogeneous social learner populations. This is most pronounced when $L=1$, and even more
so when $\pilow=0.45$ and $B=10$ (Figure~\ref{fig:payoffs}C).
In fact, social learning evolved even in parts of the parameter space where the expected homogeneous social-learning payoff was lower than the expected homogeneous
asocial-learning payoff (Figure~\ref{fig:payoffs}B, e.g., $L=1$).

@@

When payoff ambiguity and selection-set size was small ($\pilow=0.1$ and $B=2$), and lifespan was long ($L \geq 4$), asocial learning fixated as a hedge against periods of protracted environmental instability, even though expected
homogeneous social payoffs were greater than homogeneous asocial payoffs (Figure~\ref{fig:payoffTimeseries}A). Runs of environmental change strongly affected the geometric mean fitness of social learners who were weighing
outdated information too heavily. \edit{In other words, environmental change lowers social learners' mean payoffs, but social learner payoffs build back up when (blue dots, Figure~\ref{fig:payoffTimeseries}A).  Even though an homogenous population of social learners would outperform individual learners in this case, unpredictable environmental stochasticity undermines the reliability of social learning.  When lifespan was minimal ($L=1$) and environmental variability was moderate ($u\approx0.5$), the opposite, unexpected outcome could occur where social learning fixates despite the fact that an homogenous population of asocial learners would outperform an homogenous population of social learners (Figure~\ref{fig:payoffTimeseries}B). In this case, it takes relatively few consecutive generations with environmental stability to attract the population to all become social learners.}

\begin{figure}
  \caption{Average number of generations (y-axes) to fixation. 
    Note the y-axis scale varies between plots, indicating different overall 
    time to fixation for different $(\pilow, B)$ combinations.} 
  \label{fig:steps}
\centering
    \includegraphics[width=\textwidth]{Figures/stepResultsPlots.pdf}
\end{figure}

@@

\begin{figure}
  \addtocounter{figure}{-1}
    \caption{Example time series of geometric moving average (GMA) with window of 3
      of payoffs from two select uncertainty settings (see main text), broken out by
    social learners, asocial learners, and whole population, compared with the
    expected homogeneous social and asocial population payoffs. Social-learner prevalence is also plotted. Vertical lines indicate environmental change.}
    \label{fig:payoffTimeseries}
  \centering

\vspace{1em}
\hspace{-2em}

      \centering
      % \caption{$u=0.2$, $\pilow=0.1$, $B=2$, $L=4$}
      {\textsf A, Environmental variability $=$ 0.2, Low payoff $=$ 0.1, selection-set size $=$ 2, effective lifespan $=$ 4} \\
      \includegraphics[width=0.85\textwidth]{Figures/geopayoff_series/geopayseries_u=0.2-lowpayoff=0.1-nbehaviors=2-L=4.pdf} \\

    % \caption{$u=0.5$, $\pilow=0.1$, $B=4$, $L=1$}
      {\textsf B, Environmental variability $=$ 0.5, Low payoff $=$ 0.1, selection-set size $=$ 4, effective lifespan $=$ 1} \\
      \includegraphics[width=0.85\textwidth]{Figures/geopayoff_series/geopayseries_u=0.5-lowpayoff=0.1-nbehaviors=4-L=1.pdf}

\hspace{-2em}

@@

 Despite the common claim that copying others makes sense when one is uncertain, we find that different forms of uncertainty can
 either promote or impede the evolution of social learning. By disambiguating various formalizations of uncertainty in a single
 computational model, we have developed a more nuanced theoretical understanding of how uncertainty affects the evolution of social learning. %(but see \citet{healy2017}).
%We reviewed and distilled several empirical and theoretical models in the literature along four uncertainty dimensions: temporal environmental uncertainty, selection-set size, ambiguous payoffs, and effective lifespan.  
We reproduced the well-known results that the uncertainty derived from a temporally-variable environment can limit the evolution of social learning if environmental change is simply too frequent for intergenerational transmission to be of value. Importantly, we also showed that this relationship is affected by other sources of uncertainty. 
%Our model reproduced classic predictions that social learning can yield higher payoffs than individual learning when uncertainty is sufficiently limited. 
We found that larger sets of behavioral options, shorter effective lifespans,
and increased payoff ambiguity could all favor the evolution of social learning.
However, these effects were not always straightforward. Increased payoff
ambiguity and decreased effective lifespan weakened selection strength
(i.e., evolutionary uncertainty in a finite-size population). Not surprisingly,
when the consequence of one's choice does not affect payoffs much, selection
either for or against social learning is substantially weaker. More
interestingly, selection is also weaker when the ability to learn individually
is fraught with uncertainty. Our model demonstrated that social learning can act
as a scaffold for \edit{adaptive} individual learning. We endowed agents with a
biologically-plausible individual learning mechanism \edit{based on a softmax decision rule}. This
enabled agents to recover from the potential cost of receiving outdated
information. This, in turn, enabled the evolution of social learning even when
social information was likely to be outdated.

@@

Our model necessarily made simplifying assumptions. These were chosen judiciously to
address the main question of how uncertainty affects the evolution of social
learning. Two critical assumptions we made have important consequences for comparing
our results with previous findings in the literature. First, we assumed all agents
engage in individual learning. This means that individual learning is effectively
free, though potentially inaccurate. Although producing new knowledge may be costly,
many other adaptive problems take the form of essentially cost-free individual
learning. For example, it is hard not to encode association between clouds and rain,
between thirst and drinking water, or between friendship and warm laughter.
\edit{If there were an exogenous cost of individual learning, this would likely increase the social learning extinction variability across uncertainty parameter settings.
Our model operationalizes environmental variability as fully stochastic,
which forces evolution towards either a population of social or asocial learners
since environmental change is unpredictable, unlike precursor
models that allowed for persistent frequency
dependence~\citep{BoydRicherson1985,Rogers1988,Feldman1996}.} Second, we only allow for
intergenerational social learning and environmental change. Within generations,
agents can only learn individually. This assumption allows us to replicate the
classic finding that if the environment changes too quickly, socially-learned
information may become \edit{outdated. Adding horizontal social learning between
vertical/oblique learning in this model would likely expand the range of
environmental variability over which social learning would
evolve~\citep{Turner2022}, i.e., horizontal learning would increase what we have
called the social learning extinction variability. However, depending on the
setting, there may be little difference between horizontal and oblique learning, for
example if the generations are purely cultural ‘generations’ where parents, and
others of the parent's generation, may also be peers.}

@@

%We found that the 
The evolution of social learning is a complex phenomenon, dependent on many interrelated factors, yet it is widespread across a
range of taxa. Although social learning is broadly studied, few studies have taken a systematic approach to understanding the role
of uncertainty in the evolution \edit{of} social learning. Understanding the role of uncertainty in social evolution is particularly important
for understanding human evolution as evidence increasingly suggests that the genus \emph{Homo} emerged in the context of substantial
(nonstationary) environmental variability \citep{anton_etal2014, levin2015}, and our global dispersal
suggests that human groups regularly face novel environments. \edit{In our model, more numerous behavioral
options led agents to be more cognitively flexible via the softmax decision rule, and thus better able to
adjust their behavior if social learning provided outdated information. While Rogers'-style models of social
learning evolution may also predict greater social learning reliance in more challenging contexts (larger
selection-set sizes), they cannot predict that social learning should be favored for such high values of
environmental variability as we observed here.} 
By carefully identifying and operationalizing different forms of uncertainty common in social-learning models, we developed a more thorough understanding of how uncertainty can favor social learning and how different forms of uncertainty interact to favor its evolution. By explicitly modeling a key individual-level learning mechanism shared across taxa, namely the ability to adjust behavior to uncertainty and new information, we found that social learning could evolve even if it provides outdated information. %Social learning is a critical component of problem solving among social creatures. Our work identified and systematized some important forms of uncertainty and catalogued their effects on the evolution of social learning.
Our work, then, both advances the theory of social learning and could eventually be of practical use in identifying specific mechanisms underlying adaptations to uncertain environments.

@@

\paragraph{Author contributions.} JHJ, CM, PES, and MAT conceived of the project. MAT developed the model
and analyses, with contributions from CM, PES, and JHJ. MAT implemented and analyzed the model. MAT wrote and edited the initial manuscript draft, which was extensively expanded and edited by CM, PES, and JHJ.
